Stochastic blockmodeling of relational event dynamics

### 2012-10-31 docs/submission.tex
Add a few extra sentences regarding Figure 2.

One can extend the above to marked point processes where each event contains additional information.
In the context of relational events occurring among $N$ nodes, each event in the process contains both a sender $i$ and a recipient $j$ such that  $(i,j) \in \mathcal{R}$, where $\mathcal{R}$ is the \emph{risk set} is defined as the set of all possible dyadic events.
We assume each event involving dyad $(i,j) \in \mathcal{R}$ occurs as a Poisson process with intensity $\lambda_{i,j}(t|\cdot)$ that is piecewise constant and depends on the previous history of events  $\mathcal{A}_t = \{(t_m,i_m,j_m): t_m \in [0,t) \}$.  
%The likelihood of an observed event history $\mathcal{A}$ extending from time 0 to the time of the final event, $t_M$, is 
The likelihood of an observed event history $\mathcal{A}_{t_M}$ (extending from time 0 to the time of the final event and denoted $\mathcal{A}$ for convenience) is
%The likelihood of an observed event history $\mathcal{A}_{t_M}$ (also denoted $\mathcal{A}$ for convenience) is
\begin{align}
\mathcal{L}(\mathcal{A}|\theta) &= \prod_{m=1}^M \lambda_{i_m,j_m}(t_m|\cdot) \prod_{(i,j) \in \mathcal{R}}\exp\{ - (t_m - t_{m-1}) \lambda_{i,j}(t_m | \cdot) \}
\label{eqn:llk}
\end{align}

@@

% In practice we may believe an interaction between one dyad should not affect the rate of an entirely separate dyad.
As shown in Figure \ref{fig:example}, we only allow each intensity function $\lambda_{i,j}(t)$ to change following an event where $i$ was the sender or the recipient.
This is sensible in distributed settings where $i$ has limited knowledge about interactions among other actors.
In the Supplementary Material we show how this assumption also reduces the computational complexity of computing the above likelihood.

@@

We additionally perform experiments that evaluate the likelihood of future data under our model.   As our method jointly models \emph{which} dyads occur and \emph{when} they occur, any alternative method needs to also model these two aspects.  In this vein, we extend the above baselines by assuming each dyad is a Poisson process with estimated  rate $$\hat{\lambda}_{i,j}(t_m) = \frac{M}{t_M} \frac{r_{b}(i,j) + \xi}{\sum_{i,j} r_{b}(i,j) + \xi}$$, where $r_b(i,j)$ is the statistic for a given baseline (described above) and $\xi=1$ is a smoothing parameter.  
In Table ~\ref{tab:results} we compare the loglikelihood of held-out test data under our proposed model to each of these baselines.  In order to investigate the role of the number of clusters, we introduce an upper bound $K^*$ on the number of clusters during the fitting process.
Note that the model with $K*=1$ is an important baseline -- it is simply a relational event model with no groups. There are a large number of other possible baselines that one might use, but most are not directly applicable to Table 2 as they do not work with events in continuous time.

@@

%\subsubsection*{Acknowledgements}
\newpage
\newpage
\bibliographystyle{unsrt}
\bibliography{refs}
\end{document}

>>>>

One can extend the above to marked point processes where each event contains additional information.  {\color{red} In Figure \ref{fig:example}a we illustrate the case of \emph{relational events}  occurring among $N$ nodes}, where each event in the process contains both a sender $i$ and a recipient $j$ such that  $(i,j) \in \mathcal{R}$, where $\mathcal{R}$ is the \emph{risk set} is defined as the set of all possible dyadic events.
{\color{red} The study of human communication often involves such data (e.g. phone calls, online interaction, etc.) where the nodes represent people and each event represents one person speaking to another.}

As in \cite{Butts} in the proposed approach we assume each event involving dyad $(i,j) \in \mathcal{R}$ occurs as a Poisson process with intensity $\lambda_{i,j}(t|\cdot)$ that is piecewise constant and depends on the previous history of events  $\mathcal{A}_t = \{(t_m,i_m,j_m): t_m \in [0,t) \}$.  
%The likelihood of an observed event history $\mathcal{A}$ extending from time 0 to the time of the final event, $t_M$, is 
The likelihood of an observed event history $\mathcal{A}_{t_M}$ (extending from time 0 to the time of the final event and denoted $\mathcal{A}$ for convenience) is
%The likelihood of an observed event history $\mathcal{A}_{t_M}$ (also denoted $\mathcal{A}$ for convenience) is
\begin{align}
\mathcal{L}(\mathcal{A}|\theta) &= \prod_{m=1}^M \lambda_{i_m,j_m}(t_m|\cdot) \prod_{(i,j) \in \mathcal{R}}\exp\{ - (t_m - t_{m-1}) \lambda_{i,j}(t_m | \cdot) \}
\label{eqn:llk}
\end{align}

@@

% In practice we may believe an interaction between one dyad should not affect the rate of an entirely separate dyad.
As shown in Figure \ref{fig:example}, we only allow each intensity function $\lambda_{i,j}(t)$ to change following an event where $i$ was the sender or the recipient.
This is sensible in distributed settings where $i$ has limited knowledge about interactions among other actors.
% In the Supplementary Material we show how this assumption also reduces the computational complexity of computing the above likelihood.

@@

We additionally perform experiments that evaluate the likelihood of future data under our model.   As our method jointly models \emph{which} dyads occur and \emph{when} they occur, any alternative method needs to also model these two aspects.  In this vein, we extend the above baselines by assuming each dyad is a Poisson process with estimated  rate $$\hat{\lambda}_{i,j}(t_m) = \frac{M}{t_M} \frac{r_{b}(i,j) + \xi}{\sum_{i,j} r_{b}(i,j) + \xi}$$, where $r_b(i,j)$ is the statistic for a given baseline (described above) and $\xi=1$ is a smoothing parameter.  
In Table ~\ref{tab:results} we compare the loglikelihood of held-out test data under our proposed model to each of these baselines.  In order to investigate the role of the number of clusters, we introduce an upper bound $K^*$ on the number of clusters during the fitting process.
Note that the model with $K^*=1$ is an important baseline -- it is simply a relational event model with no groups. There are a large number of other possible baselines that one might use, but most are not directly applicable to Table 2 as they do not work with events in continuous time.

@@

%\subsubsection*{Acknowledgements}

\clearpage
\bibliographystyle{unsrt}
\bibliography{refs}
\end{document}
